---
title: "Calibration - Static Model"
output: html_document
date: "2024-01-08"
---


# Calibrate a 2x2 model 



See:  https://docs.google.com/spreadsheets/d/1LbwFNB81VTQSGGbJ1iwFhSimmtGjjZ_XQnqvxcTJhKA/edit?usp=sharing

N=2 Securities		
K=2 states		
		
Targets:		
R_f	5.6%	 .. per year
E[R_m-R_f]	6.0%	 .. per year
sigma[R_m-R_f]	20.0%	 .. per year
Sharpe Ratio (FYI)	0.30	 .. per year
Current S&P Price	$4,756.04	.. as of 2023.01.8


```{r}


# Load Data and few libs  
library("tidyverse")
library("tidyquant")
library("lubridate")
library("zoo")
library("gt")  # use "gt" instead of print for tables

# Bryan specific ones 
source("/Users/rout/Desktop/Dropbox/R_lib/bryan_lib.R")
source("~/Desktop/Dropbox/R_lib/fred_lib.R")  # loads api key and fred library

source("returns_lib.R")
```


- Focus on 2 state one for now

GET LIVE:
- Rf =  https://fred.stlouisfed.org/series/DGS1MO
- sigma.eRm =  https://fred.stlouisfed.org/series/VIXCLS
- P.stock = https://fred.stlouisfed.org/series/SP500

```{r}
Live.Data.Series<-tribble(
	 ~series_id, ~series_name_short, ~series_name, 
  #----------- | --------	 
  "DGS1MO",   "Rf", "Treasury Securities at 1-Month Constant Maturity",
  "SP500",   "P.stock",  "S&P 500",
  "VIXCLS",  "sigma.eRm", "CBOE Volatility Index: VIX"
)

Data<-get.fred.data(series = Live.Data.Series)

Data <- Data %>% mutate(
              # hack for now; fix later
              value=ifelse(series_name_short %in% c("Rf","sigma.eRm"),value/100,value)
            )
  
Meta<-get.fred.meta(series = Live.Data.Series)

Data.Latest <- Data %>% group_by(series_id) %>% filter(row_number()==n())
                               
```

```{r}
Get.Live <- function(
    X,
    data=Data.Latest
){
  for (i in data %>% pull(series_name_short)){
    print(i)
    x <- data %>% filter(series_name_short == i) %>% pull(value)
    X <- X %>% mutate(!!i := x)
    
    t <- data %>% filter(series_name_short == i) %>% pull(date)
    i<-paste0(i,".date")
    X <- X %>% mutate(!!i := t)
  }
  X<- X %>% relocate((ends_with("date")), .after = last_col())
  return(X)
}
```


```{r}
Economy <- list(
  K=2,   # States
  pL = 0.5,
  pH = 0.5,
  #
  Rf =   0.05,
  E.eRm = 0.06,
  sigma.eRm = 0.20,
  #
  P.bond=100.0,
  P.stock = 4756.04
) %>% as_tibble()


# Economy <- Get.Live(Economy)  # updates a Rf etc.

Economy %>% gt()

```

## Make D R P Matricies 

- hack for now
- gausian quadriture for more general examples later

NOTE:  https://css18.github.io/linear-algebra.html

```{r}

#initialize
R <-  matrix(c(1,1,1,1), nrow = 2, ncol = 2, byrow = TRUE)
D <-  R  #same
P <- c(1,1)            #matrix(c(1,1), nrow = 1, ncol = 2, byrow = TRUE)
prob <- c(.5,.5)
a<-prob # AD
m<-prob # kernel
q<-prob # rn probs

n.rf=1
n.stock=2
k.bust=1
k.boom=2

prob[k.bust]=Economy$pL
prob[k.boom]=Economy$pH

R[k.bust,n.rf]=1 + Economy$Rf 
R[k.boom,n.rf]=1 + Economy$Rf


R[k.bust,n.stock]=1 + Economy$Rf + 
        Economy$E.eRm - sqrt((1-prob[k.bust])/prob[k.bust])*Economy$sigma.eRm
R[k.boom,n.stock]=1 + Economy$Rf + 
        Economy$E.eRm + sqrt((1-prob[k.boom])/prob[k.boom])*Economy$sigma.eRm


P[n.rf] = Economy$P.bond
P[n.stock] = Economy$P.stock 

D = R * rbind(t(P),t(P))  # element by element ... just re-scales

```


- Calculate AD prices and Q and m

Theta D = I 
Theta = D^{-1}
a' = P Theta = P D^{-1}
or
a'D = P
a' = P Theta = P D^{-1}



```{r}
D.inv = solve(D) 
a =  t( P %*%  D.inv )  %>% as.vector()  # The t() so it is a row matrix 
```

- Normalizations

m = a(s) /p(s)   ... cost of $1 in state per unit of probability
b = price of certain $1 = sum(a)
q = a(s) /sum(a)

```{r}
m = a / prob   # element by element

q = a / sum(a)  # note sum(a) is the rf bond price (for payoff $1.00)

```


# Display all this in a pretty table ... 




# Eigenvectors and Eigenvalues

```{r}
X <- eigen(R)

R.eigenvalues <- X$values
R.eigenvectors <- X$vectors

#
R.equal.to.R <- R.eigenvectors %*% diag(R.eigenvalues) %*% solve(R.eigenvectors)

```

